---
title: "Session one"
author: "Mees van der Ent"
date: "10/12/2021"
output: pdf_document
---

```{r}
wd <- getwd()
data_dir <- paste(wd, "/data/", sep="")
out_dir <- paste(wd, "/out/", sep="")
```


## Chapter 1 

### 1.1: Common exercise
Create two arrays x and y and plot as coordinates in scatterplot.
```{r}
x <- c(2, 4, 7, 10)
y <- c(1, 10, 11, 18)

x
y

plot(x, y)
```

\newpage
## Chapter 2

### 2.1: Vectors
a. Create the object test1 with numbers 1.5, 0.7, 45.6.
b. Create a vector y1 with the numbers from 1 to 10.
c. Create a logical vector y2 from y1. An element of y2 should be TRUE if the corresponding element of
y1 is larger than 5.
d. How many elements from y1 has a value larger than 5?
```{r}

test1 <- c(1.5, 0.7, 45.6)
y1 <- 1:10

y2 <- y1>5
sum(y2)
```

### 2.2: creating sequences

a. Create a vector x with elements (1, 2, 3, . . . , 100)  
```{r}
x <- 1:100
x
```

b. Create a vector y with elements (0, 5, 10, 15, . . . , 500)
```{r}
y <- seq(from=0, to=500, by=5)
y
```

c. Create a vector z1 with elements (1, 1, 1, 2, 2, 2, . . . , 50, 50, 50)
```{r}
z1 <- rep(1:50, rep(3, 50))
z1
```

d. Create a vector z2 with elements (1, 2, 2, 3, 3, 3, . . . , 10)
```{r}
z2 <- rep(1:10, 1:10)
z2
```

e. Create a vector z3 with elements (1, 2, 2, 3, 4, 4, 5, 6, 6, . . . , 50, 50)
```{r}
z3 <- rep(1:50, rep(1:2, 25))
z3
```

### 2.3: Matrix creation
a. Create a vector with 100 random normal numbers and use that to generate a 10 by 10 matrix. Call this matrix mat1.
Hint: to generate a random normal vector, use the function rnorm().
```{r}
rand_dist <- rnorm(100)
mat1 <- matrix(rand_dist, nrow=10, ncol=10)
mat1
```

b. Add an extra row to mat1 with the numbers 1 to 10 which will be the new first row. Also add the row with numbers 10 to 1 which will be the last row.
```{r}
mat2 <- rbind(1:10, mat1, 10:1)
mat2
```

c. Add an extra column to the matrix obtained in step b with the number 1 to 12 (as first column).
```{r}
mat3 <- cbind(1:12, mat2)
mat3
```

### 2.4: Working with data frames
a. Install and load the package *reshape*.
```{r}
library(reshape)
```
b. Check the data description of this data frame.
```{r}
?tips
```
c. Ask for the names of the variables in this data frame.
```{r}
names(tips)
```
d. Take a subset of data tips which contains the observations from 1 until 20 and only the variables *tip*, *sex* and *day*.
```{r}
tips_subset <- tips[1:20, c("tip", "sex", "day")]
tips_subset
```


## Chapter 3
### 3.2: Importing an *Excel* file using R function
```{r}
library(readxl)
```
```{r}
titanic_file <- paste(data_dir, "titanic.xlsx", sep="")

titanic_df <- read_excel(titanic_file)
titanic_df
```

### 3.3: Export a data frame to at *xlsx* file
```{r}
library(openxlsx)

write.xlsx(airquality, file=paste(out_dir, "airquality.xlsx", sep=""))
```

### 3.4: Importing a /txt/ file using the read.table() function
```{r}
chol <- read.table(file=paste0(data_dir, "chol_R.txt"), header=TRUE)
chol
```

### 3.5: Export a data frame to a *.txt* file
```{r}
write.table(
  chol, 
  file=paste0(out_dir, "chol_export.txt"), 
  quote=FALSE,
  sep=" ",
  row.names=FALSE,
  col.names = TRUE)
```


\newpage

## Chapter 4: Functions

1. Write a function which gives the most elementary statistics for a sample x: min, median, max, mean, sd and length. Apply your function on a vector x with values from 25 to 80.

```{r}
elem_stat <- function(array){
  array_min <- min(array)
  array_max <- max(array)
  array_median <- median(array)
  array_mean <- mean(array)
  array_sd <- sd(array)
  array_length <- length(array)
  stats <- c(array_min, array_max, array_median, array_mean, array_sd, array_length)
  
  return(stats)
}

test_array <- 25:80
elem_stat(test_array)
```

2. Write a function fun1 which produces the text ‘Non-negative number’ if you apply fun1 to a positive number and ‘negative number’ if you apply fun1 to a negative number. You can make use of the ifelse function in R. Apply this function to the values 9 and -13.
```{r}
fun1 <- function(n){
  if(n>=0){
    print("Non-negative number")
  } else{
    print("Negative number")
  }
}

x <- 9; fun1(x)
x <- -13; fun1(x)
```

3. Write a function to solve an equation of second degree (ax2 + bx + c = 0).
```{r}
determinant <- function(a, b, c){
  return(b^2 - (4*a*c))
}

solve_second_deg <- function(a, b, c){
  d <- determinant(a, b, c)
  if(d>0){
    x1 <- (-b + sqrt(d)) / (2 * a)
    x2 <- (-b - sqrt(d)) / (2 * a)
    return(c(x1, x2))
  } else if(d==0){
    return((-b + sqrt(d)) / (2 * a))
  } else{
    print("determinant < 0")
  }
}

fun_second_fun <- function(x, a, b, c){
  y <- a*x^2 + b*x + c
  return(y)
}

my_second_fun <- function(x){
  return(fun_second_fun(x, -8, 6, 4))
}

print(solve_second_deg(-8, 6, 4))
plot(my_second_fun(seq(from=-2, to=2, by=0.01)), type="l")

```

