---
title: "Session one"
author: "Mees van der Ent"
date: "10/12/2021"
output: pdf_document
---

## Chapter 1 

### 1.1: Common exercise
Create two arrays x and y and plot as coordinates in scatterplot.
```{r}
x <- c(2, 4, 7, 10)
y <- c(1, 10, 11, 18)

x
y

plot(x, y)
```

\newpage
## Chapter 2

### 2.1: Vectors
a. Create the object test1 with numbers 1.5, 0.7, 45.6.
b. Create a vector y1 with the numbers from 1 to 10.
c. Create a logical vector y2 from y1. An element of y2 should be TRUE if the corresponding element of
y1 is larger than 5.
d. How many elements from y1 has a value larger than 5?
```{r}

test1 <- c(1.5, 0.7, 45.6)
y1 <- 1:10

y2 <- y1>5
sum(y2)
```

### 2.2: creating sequences

a. Create a vector x with elements (1, 2, 3, . . . , 100)  
```{r}
x <- 1:100
x
```

b. Create a vector y with elements (0, 5, 10, 15, . . . , 500)
```{r}
y <- seq(from=0, to=500, by=5)
y
```

c. Create a vector z1 with elements (1, 1, 1, 2, 2, 2, . . . , 50, 50, 50)
```{r}
z1 <- rep(1:50, rep(3, 50))
z1
```

d. Create a vector z2 with elements (1, 2, 2, 3, 3, 3, . . . , 10)
```{r}
z2 <- rep(1:10, 1:10)
z2
```

e. Create a vector z3 with elements (1, 2, 2, 3, 4, 4, 5, 6, 6, . . . , 50, 50)
```{r}
z3 <- rep(1:50, rep(1:2, 25))
z3
```

### 2.3: Matrix creation
a. Create a vector with 100 random normal numbers and use that to generate a 10 by 10 matrix. Call this matrix mat1.
Hint: to generate a random normal vector, use the function rnorm().
```{r}
rand_dist <- rnorm(100)
mat1 <- matrix(rand_dist, nrow=10, ncol=10)
mat1
```

b. Add an extra row to mat1 with the numbers 1 to 10 which will be the new first row. Also add the row with numbers 10 to 1 which will be the last row.
```{r}
mat2 <- rbind(1:10, mat1, 10:1)
mat2
```

c. Add an extra column to the matrix obtained in step b with the number 1 to 12 (as first column).
```{r}
mat3 <- cbind(1:12, mat2)
mat3
```

### 2.4: Working with data frames
a. Install and load the package *reshape*.
```{r}
library(reshape)
```
b. Check the data description of this data frame.
```{r}
?tips
```
c. Ask for the names of the variables in this data frame.
```{r}
names(tips)
```
d. Take a subset of data tips which contains the observations from 1 until 20 and only the variables *tip*, *sex* and *day*.
```{r}
tips_subset <- tips[1:20, c("tip", "sex", "day")]
tips_subset
```


